\documentclass{report}
\title{latex-notes}

\usepackage[utf8]{inputenc}
\usepackage{color}
\usepackage{hyperref}
\usepackage{tabto}

\begin{document}
\pagenumbering{gobble}
\tableofcontents


\part{TODO}
\begin{itemize}
	\item check how to write a tilde and update the list of 'Special Characters'
	\item how to use maketitle, 'Preamble Commands. \textbackslash maketitle.'
	\item \textbackslash setcounter
	\item crossreferences \textbackslash ref an \textbackslash label
	\item KOMA script
\end{itemize}

\part{Preamble Commands}
\section{\textbackslash documentclass{book, article...}}
\section{\textbackslash usepackage{…}}
\section{\textbackslash maketitle}




\part{Text Formatting}



\chapter{Lists}


\section{Bullet}
\begin{verbatim}
\begin{itemized}
    \item 
    ...
\end{itemized}
\end{verbatim}


\section{Numbered}	
\begin{verbatim}
\begin{enumerate}
    \item 
    ...
\end{enumerate}
\end{verbatim}


\part{no-title}



\chapter{Text Layout}


\section{Tabulators}

\subsection{tabto}
\begin{verbatim}
\usepackage{tabto}
...
left \tab right
\end{verbatim}
produces:\\
\\
left \tab right

\section{paragraph} - can be done by making an empty line:

\begin{verbatim}
  bla bla

  bla bla
\end{verbatim}

1st paragraph. alalala alalala alalala lalalala lalalalala alalalal alalala alalala alalala alalal alalala .

2nd paragraph alalala alalala alalala alalala alalala 

\section{arguments}
Two types:
\begin{itemize}
	\item \textbf{mandatory}
	\begin{itemize}
		\item 	enclosed within a pair of \textbf{curly brackets}.
		\item each argument is between a separate pair of brackets.
	\end{itemize}

	\item \textbf{optional}\\
	\textcolor{red}{\textbf{all} enclosed within a pair of \textbf{single pair of square brackets}}.
\end{itemize}

\section{An Empty Space after no-arg Commands}
No-arg commands 'consume' empty spaces that follow. to avoid it"
\begin{itemize}
	\item add empty argument list - \{\}, or
	\item put inside curly brackets
\end{itemize} 
\begin{verbatim}
\LaTeX{} {\LaTeX} \LaTeX
\end{verbatim}
outputs:\\
\LaTeX{} {\LaTeX} \LaTeX,
\begin{verbatim}
\LaTeX \LaTeX
\end{verbatim}
outputs no space between:\\
\LaTeX \LaTeX

\section{Special Characters}
\begin{itemize}
	\item  \#{}  \tab \textcolor{red}{\textbf{\textbackslash \#}}
	\item \${} \tab \textbf{\textcolor{red}{ \textbackslash \$}}
	\item \%{} \tab \textbf{\textcolor{red}{\textbackslash \%}}
	\item \^{} \tab \textcolor{red}{\textbf{\textbackslash \^{} \{ \}}}
	\item \&{}\tab \textbf{\textcolor{red}{\textbackslash \&}}
	\item \_{} \tab \textbf{\textcolor{red}{\textbackslash \_}}
	\item \{{} \tab \textbf{\textcolor{red}{\textbackslash \{}}
	\item \}{} \tab \textbf{\textcolor{red}{\textbackslash \}}}
	\item  \~{} \tab \textbf{\textcolor{red}{\textbackslash \~{}\{{}\}{}}} 
	\item \textbackslash \tab \textbf{\textcolor{red}{\textbackslash textbackslash}}
\end{itemize}

\part{Environments}
\chapter{Page Layouts}
\section{quote}
\begin{verbatim}
\begin{quote}
   This is a quote. This is a quote. This is a quote. This is a quote. ...
   This is a quote. This is a quote. This is a quote. This is a quote.
\end{quote}
\end{verbatim}
… adds extra space to both margins:\\
A quote follows:
\begin{quote}
	This is a quote. This is a quote. This is a quote. This is a quote. This is a quote.
	This is a quote. This is a quote. This is a quote. This is a quote. This is a quote.
	This is a quote. This is a quote. This is a quote. This is a quote. This is a quote.
\end{quote}

\end{document}