\documentclass{report}
\title{Track Notes}

\usepackage[utf8]{inputenc}
\usepackage{color}
\usepackage{hyperref}
\usepackage{breakurl}

\usepackage{tabto}

\begin{document}
\tableofcontents

\chapter{Installation}
\section{Windows}
\url{https://trac.edgewall.org/wiki/TracInstall}

\begin{verbatim}
easy_install Trac
\end{verbatim}

\chapter{Setup}
\url{http://web.archive.org/web/20100324230430/http://how-to-solutions.com/
how-to-install-trac-on-windows.html}

\begin{verbatim}
trac-admin C:\project_dir initenv
\end{verbatim}
 
In Windows open up the project folder and go into the \textit{conf} sub-directory. 
There's a file called trac.ini - open it in an editor like Notepad. You need to copy 
the following lines below and paste them into Notepad above the \textit{header\_logo} section.

\begin{verbatim}
[components]
acct_mgr.admin.accountmanageradminpage = enabled
acct_mgr.api.accountmanager = enabled
acct_mgr.db.sessionstore = enabled
acct_mgr.htfile.htdigeststore = enabled
acct_mgr.pwhash.htdigesthashmethod = enabled
acct_mgr.web_ui.accountmodule = enabled
acct_mgr.web_ui.loginmodule = enabled
acct_mgr.web_ui.registrationmodule = enabled
trac.web.auth.loginmodule = disabled
\end{verbatim}

Go to project\_dir and create a new text file called passwd.txt - leave the file blank.

Lastly let's give all anonymous users (those who aren't logged in) admin access. In the same 
CLI window type the following and hit Enter.
\begin{verbatim}
trac-admin C:\project_dir permission add anonymous TRAC_ADMIN
\end{verbatim}

\textbf{\textcolor{red}{Install AccountManager plugin now!}} - see section below.

In CLI type the following and hit Enter.
\begin{verbatim}
tracd -p 8000 E:\project
\end{verbatim}

Open up browser and paste:
\begin{verbatim}
http://localhost:8000/. 
\end{verbatim}

In the top right there's a tab called Admin - click on it. On the menu on the right
click on \textit{Configuration} and when the page loads select \textit{HtDigestStore}
where the filename is $C:\backslash project\backslash passwd.txt$ and the realm is 'trac'.

Below \textit{Configuration} is \textit{Users} - click on it. Create yourself 
an admin account. Now using the menu go to \textit{Permissions}. On the right 
of that page you will see the \textit{Grant Permission} section. Use the subject 
field for the username you gave yourself on the Users page and select the \textit{TRAC\_ADMIN} as
\textit{Action}. Click Add and you will see your username in the middle column of the page with 
\textit{TRAC\_ADMIN} next to it. For the subject anonymous click on the \textit{TRAC\_ADMIN} 
so it's ticked and then click on the \textit{Remove} selected items button.

\section{AccountManager plugin}
\begin{enumerate}
  \item download from \url{https://trac-hacks.org/wiki/AccountManagerPlugin}
  \item unpack
  \item run track server:
  \begin{verbatim}
  tracd -p 8000 projecf_dir
  \end{verbatim}
  \item go to admin, plugins (left pane)
  \item install from egg buton. egg file is in \textit{dist}
  subdir of unpacked folder
\end{enumerate}

\end{document}